A Framework for Modelling Thin Objects: \\ from Meshing to Shell Finite Element Method

### 2010-03-10 MICCAI2010/paper.tex
abstract + qqs modifs
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@93 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\begin{abstract}

\end{abstract}

@@

Shell theory allows the modelling of  structure  deformations when the thickness is small compared to its other dimensions~\cite{Liu03}. The key idea is to model the physical shell as if it were a surface of zero thickness, but endowed with mechanical properties in the form of elastic resistance to stretching, twisting, bending and shearing forces.
Rather than resorting to shell theory, previous work on medical simulation often relies on mass-spring models (see \ref{Hammer08} \CD{voir fichier bib isbms} for instance).
%Rather than resorting to shell theory which involves the most complex formulations in continuum mechanics, previous works have often relied on discrete formulations. Early approaches only considered in-plane deformation, and often relied on mass-spring models. More recent works have considered adding bending through angular springs. For instance Wang et al.~\cite{Wang07} successfully used a network of linear and angular springs to describe bending and twisting of catheters and guidewires in an interventional radiology simulator. \CD{C'est pas des shells si?} 
Yet, such models are limited in their ability to describe certain behaviour, as they do not rely on continuum mechanics: it is difficult to derive spring stiffness (in particular for angular springs) from elastic properties (Young's modulus and Poisson's ratio).
The use of fast shell based model for computer graphics and interactive simulation was recently introduced with the work of Choi~\textit{et al.}~\cite{Choi}. The model relies on simplified energy functions and precomputed modal analysis. It allows for fast and visually realistic results.
We propose to rely an a similar approach but with more accuracy in order to be applicable to medical simulation. Our model is not based on modal analysis but uses a corotational formulation and polynomial shape functions. In \cite{Comas2010ISBMS}, we present and apply this shell model for a flat deformable implant.

@@

\subsubsection{Mechanical interactions with the curved surface of shells}
\label{sec:interactions}
But the practical interest of modelling complex behaviours such as bending and twisting would remain fairly low for medical simulation if contacts were not handled properly. In our case the difficulty comes from different sources. First the collision detection must be carried out with the curved surface of shell elements as opposed to the classic detection on plane triangles. Then forces applied to a given triangle need to be distributed between linear forces and torques onto its three vertices.  As we will see, the same polynomial interpolation function \ref{eq-deflection} chosen to compute the bending energy in our FEM formulation is also used to capture the interactions between the curved surface and other objects.

>>>>

\begin{abstract}
This paper presents a new modeling technique for the deformation of thin anatomical structures like membrane and hollow organs. 
We show that the behavior of this type of tissue can be abstracted with a 2D modeling of their surface elastic resistance using shell theory. 
In order to apply the shell theory in the particular context of medical simulation, our method propose to base the geometrical reconstruction of the organ on the shape functions of the shell element. 
Moreover, we also use these continuous shape functions to handle the contacts and the interactions with other types of deformable tissues. The technique is illustrated using several examples including the simulation of angioplasty procedure.
\end{abstract}

@@

Shell theory allows the modelling of  structure  deformations when the thickness is small compared to its other dimensions~\cite{Liu03}. The key idea is to model the physical shell as if it were a surface of zero thickness, but endowed with mechanical properties in the form of elastic resistance to stretching, bending and shearing forces.
Rather than resorting to shell theory, previous work on medical simulation often relies on mass-spring models (see \ref{Hammer08} \CD{voir fichier bib isbms} for instance).
%Rather than resorting to shell theory which involves the most complex formulations in continuum mechanics, previous works have often relied on discrete formulations. Early approaches only considered in-plane deformation, and often relied on mass-spring models. More recent works have considered adding bending through angular springs. For instance Wang et al.~\cite{Wang07} successfully used a network of linear and angular springs to describe bending and twisting of catheters and guidewires in an interventional radiology simulator. \CD{C'est pas des shells si?} 
Yet, such models are limited in their ability to describe certain behaviour, as they do not rely on continuum mechanics: it is difficult to derive spring stiffness (in particular for angular springs) from elastic properties (Young's modulus and Poisson's ratio).
The use of fast shell based model for computer graphics and interactive simulation was recently introduced with the work of Choi~\textit{et al.}~\cite{Choi}. The model relies on simplified energy functions and precomputed modal analysis. It allows for fast and visually realistic results.
We propose to rely an a similar approach but with more accuracy in order to be applicable to medical simulation. Our model is not based on modal analysis but uses a corotational formulation and polynomial shape functions. In \cite{Comas2010ISBMS}, we present and apply this shell model for a flat deformable implant.

@@

\subsubsection{Mechanical interactions with the curved surface of shells}
\label{sec:interactions}
But the practical interest of modelling complex behaviours such as bending and twisting would remain fairly low for medical simulation if contacts and constraints were not handled properly. In our case the difficulty comes from different sources. First the collision detection must be carried out with the curved surface of shell elements as opposed to the classic detection on plane triangles. Then forces applied to a given triangle need to be distributed between linear forces and torques onto its three vertices.  As we will see, the same polynomial interpolation function \ref{eq-deflection} chosen to compute the bending energy in our FEM formulation is also used to capture the interactions between the curved surface and other objects.
